---
title: "GSP"
author: "David_M"
date: "15/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Title:  Generalized Sequence Pattern (GSP)
File:   DMR_06_02_GSP.R
Course: Data Mining with R

# INSTALL AND LOAD PACKAGES ################################

### Install pacman if you don't have it (uncomment next line)
### install.packages("pacman")

## Install and/or load packages with pacman
```{r}
pacman::p_load(  # Use p_load function from pacman
  datasets,      # R's built-in sample datasets
  magrittr,      # Pipes
  pacman,        # Load/unload packages
  rio,           # Import/export data
  tidyverse,     # So many reasons
  readxl,        # Read Xlsx File from windows
  plyr,
  dplyr,
  arules,     # For analyzing transaction data and patterns (frequent itemsets and association rules)
  arulesViz,  # visualization for association rules and item-sets
  ggplot2,    # Create graphics and charts
  lubridate,  # R package that makes it easier to work with dates and times
  RColorBrewer,
  recommenderlab
)
```

# SET RANDOM SEED ##########################################

# Set random seed for reproducibility in processes like
# sampling and splitting the data
```{r}
set.seed(1)  # You can use any number here
```

# LOAD AND PREPARE DATA ####################################

For our examples in this course, we'll use datasets from the Machine Learning Repository at the University of
California, Irvine (UCI). Here is a link to the repository's main page:
browseURL("https://archive.ics.uci.edu/ml/machine-learning-databases/00514/")


```{r}
setwd("E:\\UPWORK-2020\\Sequence_Mining_Proj\\Online_Trading")
```



```{r, message=FALSE,warning=FALSE}
# Load Data
#library("readxl")
library(data.table)
#library(knitr)

# xlsx files
retail_data <- read_excel("Online_Retail.xlsx", trim_ws = TRUE)
```

# Column Names in Data
```{r}
colnames(retail_data)
```

#Data structure
```{r}
str(retail_data)
```


```{r}
summary(retail_data)
```


```{r}
#complete.cases(data) will return a logical vector indicating which rows have no missing values. Then use the vector to get only rows that are complete using retail[,].
retail_data <- retail_data[complete.cases(retail_data), ]
```


```{r}
# Using mutate function from dplyr package to convert Description column to factor column. 
# as.factor converts column to factor column. %>% is an operator with which you may pipe values to another function or expression
retail_data %>% mutate(Description = as.factor(Description))
```


```{r, results='hide', message=FALSE,warning=FALSE}

retail_data %>% mutate(Country = as.factor(Country))

#Converts character data to date. Store InvoiceDate as date in new variable
retail_data$Date <- as.Date(retail_data$InvoiceDate)
#Extract time from InvoiceDate and store in another variable
TransTime<- format(retail_data$InvoiceDate,"%H:%M:%S")

#Convert and edit InvoiceNo into numeric
InvoiceNo <- as.numeric(as.character(retail_data$InvoiceNo))
```


```{r}
#Bind new columns TransTime and InvoiceNo into dataframe retail
cbind(retail_data,TransTime)
```



```{r}

cbind(retail_data,InvoiceNo)
```

# Removing Cancelled Orders. All Invoices starting with letter "C" imply orders were cancelled
```{r}
## if the InvoiceNo starts with letter 'C', it indicates a cancellation
retail_data %>% 
  filter(grepl("C", retail_data$InvoiceNo)) %>% 
  dplyr::summarise(Total = n())
```

```{r}
## Cancellations are not needed for the analysis so they can be removed
retail_data  <- retail_data %>% 
  filter(!grepl("C", retail_data$InvoiceNo))
```


# Filtering out Negative Quantities
```{r, message=FALSE,warning=FALSE}
## filtering by non-positive Quantity, Description shows manually entered adjustments codes. 
retail_data %>% 
  filter(Quantity <= 0) %>% 
  group_by(Description, UnitPrice) %>% 
  dplyr::summarise(count =n()) %>%
  arrange(desc(count)) %>% 
  ungroup()
```


# Removing Non - Product CODES
```{r, message=FALSE,warning=FALSE}
## There are a handful of non-product related codes - creating a filter that can be updated in the future.
stc <- c('AMAZONFEE', 'BANK CHARGES', 'C2', 'DCGSSBOY', 'DCGSSGIRL',
         'DOT', 'gift_0001_', 'PADS', 'POST')

## Summary
retail_data %>%  
  filter(grepl(paste(stc, collapse="|"), StockCode))  %>% 
  group_by(StockCode, Description) %>% 
  dplyr::summarise(count =n()) %>%
  arrange(desc(count)) %>% 
  ungroup()
```


# Cleaning up Description Column
```{r}
## Additional adjustment codes to remove - again creating a filter that can be updated in the future.
descr <- c( "check", "check?", "?", "??", "damaged", "found", 
            "adjustment", "Amazon", "AMAZON", "amazon adjust", 
            "Amazon Adjustment", "amazon sales", "Found", "FOUND",
            "found box", "Found by jackie ", "Found in w/hse", "dotcom",
            "dotcom adjust", "allocate stock for dotcom orders ta", "FBA",
            "Dotcomgiftshop Gift Voucher £100.00", "on cargo order",
            "wrongly sold (22719) barcode", "wrongly marked 23343",
            "dotcomstock", "rcvd be air temp fix for dotcom sit", "Manual",
            "John Lewis", "had been put aside", "for online retail orders",  
            "taig adjust", "amazon", "incorrectly credited C550456 see 47",
            "returned", "wrongly coded 20713", "came coded as 20713", 
            "add stock to allocate online orders", "Adjust bad debt",
            "alan hodge cant mamage this section", "website fixed",
            "did  a credit  and did not tick ret", "michel oops",
            "incorrectly credited C550456 see 47", "mailout", "test",
            "Sale error",  "Lighthouse Trading zero invc incorr", "SAMPLES",
            "Marked as 23343", "wrongly coded 23343","Adjustment", 
            "rcvd be air temp fix for dotcom sit", "Had been put aside."
)

## Filtering out the unwanted entries.
retail_data <- retail_data %>% 
  filter(!Description %in% descr)

## There are also some 600 NAs in _Description_. 
sum(is.na(retail_data$Description))
```


```{r}
#get a glimpse of your data
glimpse(retail_data)
```

# EXPLORE DATA #############################################


# EXPLORATORY DATA ANALYSIS
Most Popular Items
```{r, message=FALSE,warning=FALSE}
retail_data %>% 
  group_by(Description) %>% 
  dplyr::summarize(count = n()) %>% 
  mutate(pct=(count/sum(count))*100) %>% 
  arrange(desc(pct)) %>% 
  ungroup() %>% 
  top_n(10, wt=pct)
```


```{r}
retail_data %>% 
  group_by(Description) %>% 
  dplyr::summarize(count = n()) %>% 
  top_n(10, wt = count) %>%
  arrange(desc(count)) %>% 
  ggplot(aes(x = reorder(Description, count), y = count))+
  geom_bar(stat = "identity", fill = "royalblue", colour = "blue") +
  labs(x = "", y = "Top 10 Best Sellers", title = "Most Ordered Products") +
  coord_flip() +
  theme_grey(base_size = 12)
```


# Average Number of Items per Purchase
```{r}
retail_data %>% 
  group_by(InvoiceNo) %>% 
  dplyr::summarise(n = mean(Quantity)) %>%
  ggplot(aes(x=n)) +
  geom_histogram(bins = 100000,fill = "purple",colour = "black") + 
  coord_cartesian(xlim=c(0,100)) +
  scale_x_continuous(breaks=seq(0,100,10)) +
  labs(x = "Average Number of Items per Purchase", y = "") +
  theme_grey(base_size = 14)
```

# Saving File For Future Sequence Pattern Modeling
 
```{r}
# Create a File and store for Modeling with other Algorithms
save_data <- retail_data[ , c("CustomerID", "Date", "InvoiceNo", "Description")]

```


```{r}
write.csv(save_data,'hmm_spade_transaction_data.csv', row.names=FALSE)
```


# Take random subsample to save time (if needed)
```{r}
retail_data %<>% sample_n(80000)
```

```{r}
head(retail_data)
```


# We now convert Dataframe into Transaction Format, Also known as Singles Format (All items bought together are placed in one row)

```{r}
#library(plyr)
#ddply(dataframe, variables_to_be_used_to_split_data_frame, function_to_be_applied)
transactionData <- ddply(retail_data,c("InvoiceNo","Date", "CustomerID"),
                       function(df1)paste(df1$Description,
                       collapse = ","))
#The R function paste() concatenates vectors to character and separated results using collapse=[any optional charcater string ]. Here ',' is used
```


```{r}
# Create new column of number of count ot items for each transaction

transactionData$counts <- lengths(strsplit(transactionData$V1, ","))

transactionData
```



# Next, as InvoiceNo and Date will not be of any use in the rule mining, you can set them to NULL
```{r}
#set column InvoiceNo of dataframe transactionData  
transactionData$InvoiceNo <- NULL
#set column Date of dataframe transactionData
#transactionData$Date <- NULL
#Rename column to items
colnames(transactionData) <- c("Date", "CustomerID", "items", "counts")
#Show Dataframe transactionData
transactionData
```

# Write the transaction data into a write.csv format

```{r}
write.csv(transactionData,"market_basket_transactions.csv", quote = FALSE, row.names = FALSE)
```

# Read Transaction File
```{r, message=FALSE,warning=FALSE}
transaction_matrix <- read.transactions('market_basket_transactions.csv', format = 'basket', sep=',')
```

```{r}
transaction_matrix
```

Note: There are 18472  transactions (rows) and 12477  items (columns). Note that 12477  is the product descriptions involved in the dataset and 18472 transactions are collections of these items.



```{r}
summary(transaction_matrix)
```


```{r}
# Create an item frequency plot for the top 20 items
if (!require("RColorBrewer")) {
  # install color package of R
install.packages("RColorBrewer")
#include library RColorBrewer
library(RColorBrewer)
}

par(mar = rep(2, 4))
itemFrequencyPlot(transaction_matrix,
                  topN=20,type="absolute",
                  col=brewer.pal(8,'Pastel1'), main="Absolute Item Frequency Plot")
```


```{r}
graphics.off()
#par(mar=c(1,1,1,1))
par(mar = rep(2, 4))
itemFrequencyPlot(transaction_matrix,
                  topN=20,
                  type="relative",
                  col=brewer.pal(8,'Pastel2'),main="Relative Item Frequency Plot")

```


# Generating Rules using the APRIORI algorithm

```{r}
# Min Support as 0.001, confidence as 0.8.
association.rules <- apriori(transaction_matrix, parameter = list(supp=0.001, conf=0.8,maxlen=10))
```

```{r}
inspect(association.rules[1:10])
```

# Limiting the number and size of rules.
```{r}
shorter.association.rules <- apriori(transaction_matrix, parameter = list(supp=0.001, conf=0.8,maxlen=3))
```

# Removing redundant rules. These rules are subset of larger rules and there not neccessary.

```{r}
subset.rules <- which(colSums(is.subset(association.rules, association.rules)) > 1) # get subset rules in vector
length(subset.rules)  #> 106640
```


```{r}
subset.association.rules. <- association.rules[-subset.rules] # remove subset rules.
```


# GSP Algorithm

 - Generating Candidates and Pruning








```{r}
#write.csv(retail_data,'transaction_data.csv', row.names=FALSE)
```


# SPLIT DATA ###############################################

# Split data into training (trn) and testing (tst) sets
```{r}
#trn <- clean_data %>% sample_frac(.70)  # 70% in training data
#tst <- clean_data %>% anti_join(trn)    # Rest in testing data

```




# MODEL DATA ###############################################

**Creating The Data Matrix**
```{r}
## Filtering by an order number which contains the same stock code more than once
retail_data %>% 
  filter(InvoiceNo == 557886 & StockCode == 22436) %>% 
  select(InvoiceNo, StockCode, Quantity, UnitPrice, CustomerID)
```

# Remove Duplicates
```{r}
retail_data <- retail_data %>% 
## Create unique identifier
    mutate(InNo_Desc = paste(InvoiceNo, Description, sep = ' ')) 
## Filter out duplicates and drop unique identifier
    retail_data <- retail_data[!duplicated(retail_data$InNo_Desc), ] %>% 
    select(-InNo_Desc)
```


# Creating The Matrix
```{r}
ratings_matrix <- retail_data %>%
  
## Select only needed variables
  select(InvoiceNo, Description) %>% 

  ## Add a column of 1s
  mutate(value = 1) %>%

  ## Spread into user-item format
  spread(Description, value, fill = 0) %>%
  select(-InvoiceNo) %>%

  ## Convert to matrix
  as.matrix()

  ## Convert to recommenderlab class 'binaryRatingsMatrix'
  ratings_matrix <- as(ratings_matrix, "binaryRatingMatrix")
```


```{r}
ratings_matrix
```



# Model Evaluation Scheme and Validation

```{r, message= FALSE, warning= FALSE}
library("recommenderlab")
scheme <- ratings_matrix %>% 
  evaluationScheme(method = "cross",
                   k      = 5, 
                   train  = 0.8,  
                   given  = -1)
```

# Algorithm List
```{r}
algorithms <- list(
  "association rules" = list(name  = "AR", 
                        param = list(supp = 0.01, conf = 0.01)),
  "random items"      = list(name  = "RANDOM",  param = NULL),
  "popular items"     = list(name  = "POPULAR", param = NULL),
  "item-based CF"     = list(name  = "IBCF", param = list(k = 5)),
  "user-based CF"     = list(name  = "UBCF", 
                        param = list(method = "Cosine", nn = 500))
                   )
```


# Estimating The Models
```{r}
results <- recommenderlab::evaluate(scheme, 
                                    algorithms, 
                                    type  = "topNList", 
                                    n     = c(1, 3, 5, 10, 15, 20)
                                    )
```


# Visualizing and Evaluating The Models
```{r}
## Pull into a list all confusion matrix information for one model 
tmp <- results$`user-based CF` %>%
  getConfusionMatrix()  %>%  
  as.list() 

## Calculate average value of 5 cross-validation rounds 
  as.data.frame( Reduce("+",tmp) / length(tmp)) %>% 

    ## Add a column to mark the number of recommendations calculated
  mutate(n = c(1, 3, 5, 10, 15, 20)) %>%

    ## Select only columns needed and sorting out order 
  select('n', 'precision', 'recall', 'TPR', 'FPR')
```


```{r}
avg_conf_matr <- function(results) {
  tmp <- results %>%
    getConfusionMatrix()  %>%  
    as.list() 
    as.data.frame(Reduce("+",tmp) / length(tmp)) %>% 
    mutate(n = c(1, 3, 5, 10, 15, 20)) %>%
    select('n', 'precision', 'recall', 'TPR', 'FPR') 
}
    
## Using map() to iterate function across all models
results_tbl <- results %>%
  map(avg_conf_matr) %>% 

## Turning into an unnested tibble
  enframe() %>%

## Unnesting to have all variables on same level
unnest()

results_tbl
```


# ROC Curve
```{r}
results_tbl %>%
  ggplot(aes(FPR, TPR, 
             colour = fct_reorder2(as.factor(name), 
                      FPR, TPR))) +
  geom_line() +
  geom_label(aes(label = n))  +
  labs(title = "ROC curves", colour = "Model") +
  theme_grey(base_size = 14)
```


# Precision - Recal Curve
```{r}
results_tbl %>%
  ggplot(aes(recall, precision, 
             colour = fct_reorder2(as.factor(name),  
                      precision, recall))) +
  geom_line() +
  geom_label(aes(label = n))  +
  labs(title = "Precision-Recall curves", colour = "Model") +
  theme_grey(base_size = 14)
```


# Test Prediction
```{r}
## Create a made-up order with a string containing 6 products selected at random.
customer_order <- c("GREEN REGENCY TEACUP AND SAUCER",
                    "SET OF 3 BUTTERFLY COOKIE CUTTERS",
                    "JAM MAKING SET WITH JARS",
                    "SET OF TEA COFFEE SUGAR TINS PANTRY",
                    "SET OF 4 PANTRY JELLY MOULDS")



## Put string in a format that recommenderlab accepts.
new_order_rat_matrix <- retail_data %>% 
  select(Description) %>%
  
## Select item descriptions from retail dataset
  unique() %>% 
  mutate(value = as.numeric(Description %in% customer_order)) %>% 
  
## Add a 'value' column
  spread(key = Description, value = value) %>% 
  
## Spread into sparse matrix format
  as.matrix() 
  
## Change to a matrix
new_order_rat_matrix <-   as(new_order_rat_matrix, "binaryRatingMatrix") # Convert to recommenderlab class 'binaryRatingsMatrix'


## Create a `Recommender`
recomm <- Recommender(getData(scheme, 'train'), 
                      method = "IBCF",   
                      param = list(k = 5))


## Pass the `Recommender` and the made-up order to the `predict` function to create 
## A top 10 recommendation list for the new customer.
pred <- predict(recomm, 
                newdata = new_order_rat_matrix, 
                n       = 10)


## Inspect this pediction as a list
as(pred, 'list')
```


# CLEAN UP #################################################

# Clear data
```{r}
rm(list = ls())  # Removes all objects from the environment
```


# Clear packages
```{r}
detach("package:datasets", unload = T)  # For base packages
p_unload(all)  # Remove all contributed packages
```


# Clear plots
```{r}
graphics.off()  # Clears plots, closes all graphics devices
```


# Clear console
```{r}
cat("\014")  # Mimics ctrl+L
```

# Clear R
#   You may want to use Session > Restart R, as well, which 
#   resets changed options, relative paths, dependencies, 
#   and so on to let you start with a clean slate

# Clear mind :)
